---
title: "Intro to Cipher"
author: "Nicole LaPointe Jameson"
date: "April 18, 2016"
output: html_document
---
#Baseline Cipher Solving 

##Comparison Text
Right now, it is the book, `War and Peace`, but maybe dictionary will be better?
```{r}
#Input a baseline text fille to decode
reference <- readLines("C:\\Users\\NMLJ\\Documents\\GitHub\\FinalProject-p5-team6\\data\\warandpeace.txt")

#Translate Characters into character vectors
reference <- toupper(reference) 

```

##Probability Matrix
This produces 27 by 27 matrix (for all letters and special characters) of  the number of times  **2** letters following eachother.
*Current problem: Single-letter and double-letter words are not very accurate so far (IE: A, I, So, To)

```{r}
#Create the Matrix
prob.mat <- matrix(0,27,27)
rownames(prob.mat)=colnames(prob.mat)=c(toupper(letters),"")
lastletter=""

#The Loop to compute Log-likelihood probability
for (ln in 1:length(reference)) {
  if (ln %% 1000 ==0) {cat("Line",ln,"\n")}
  for (pos in 1:nchar(reference[ln])) {
    curletter<- substring(reference[ln],pos,pos)
    if (curletter %in% toupper(letters)) {
      prob.mat[rownames(prob.mat)==lastletter,
                colnames(prob.mat)==curletter]=
        prob.mat[rownames(prob.mat)==lastletter,
                  colnames(prob.mat)==curletter]+1
      lastletter=curletter
    } else {
      if (lastletter!="") {
        prob.mat[rownames(prob.mat)==lastletter,27]=
          prob.mat[rownames(prob.mat)==lastletter,27]+1
        lastletter=""
      }
    }
  }
  curletter=""
  if (lastletter!="") {
    prob.mat[rownames(prob.mat)==lastletter,27]=
      prob.mat[rownames(prob.mat)==lastletter,27]+1
  }
  lastletter=""
}

probability.matrix <- sweep(prob.mat+1,1,rowSums(prob.mat+1),FUN="/")

```

##Decoding Methods
This function decode() decodes the text based on the text mapping position (So based on count of how many times letters followed eachother)
```{r}
decode <- function(mapping,coded) {
  coded=toupper(coded)
  decoded=coded
  for (i in 1:nchar(coded)) {
    if (substring(coded,i,i) %in% toupper(letters)) {
      substring(decoded,i,i)=toupper(letters[mapping==substring(coded,i,i)])
    }
  }
  decoded
}

```

##Log-Likelihood Function
This function decodes based on log-likelihood of letters following eachother
```{r}
log.prob <- function(mapping,decoded) {
  logprob=0
  
  lastletter=""
  for (i in 1:nchar(decoded)) {
    curletter=substring(decoded,i,i)
    if (curletter %in% toupper(letters)) {
      logprob=logprob+log(probability.matrix[rownames(prob.mat)==lastletter,
                                         colnames(prob.mat)==curletter])
      lastletter=curletter
    } else {
      if (lastletter!="") {
        logprob=logprob+log(probability.matrix[rownames(prob.mat)==lastletter,27])
        lastletter=""
      }
    }
  }
  
  if (lastletter!="") {
    logprob=logprob+log(probability.matrix[rownames(prob.mat)==lastletter,27])
    lastletter=""
  }
  logprob
}


```

#Baseline Function
##Scrambling Methods
```{r}
#SRS of letters scramble
##Correct text (Input as character spring)
correctTxt <- "Hello Project Five Team Six! I Hope this code works."


coded <- decode(sample(toupper(letters)),correctTxt) # randomly scramble the text
coded
```

##Decoding Execution: Baseline Mapping Text (Run this to get decoded options)
```{r}
mapping <- sample(toupper(letters)) # initialize a random mapping
i=2 #Can change these values
iters=1000 #Can change these vales
cur.decode <- decode(mapping,coded)
cur.loglike <- log.prob(mapping,cur.decode)
max.loglike <- cur.loglike
max.decode <- cur.decode

while (i<=iters) {
  proposal <- sample(1:26,2) # select 2 letters to switch
  prop.mapping=mapping
  prop.mapping[proposal[1]]=mapping[proposal[2]]
  prop.mapping[proposal[2]]=mapping[proposal[1]]
  
  prop.decode=decode(prop.mapping,coded)
  prop.loglike=log.prob(prop.mapping,prop.decode)
  
  if (runif(1)<exp(prop.loglike-cur.loglike)) {
    mapping=prop.mapping
    cur.decode=prop.decode
    cur.loglike=prop.loglike
    
    if (cur.loglike>max.loglike) {
      max.loglike=cur.loglike
      max.decode=cur.decode
    }
    
    cat(i,cur.decode,"\n")
    i=i+1
  }
}

```

#Output Options
```{r}
#Regular Decode
cur.decode

#Log Likelihood 
cur.loglike

#Max Log Likelihood
max.loglike

#Max LL of Decode
max.decode

#

```


```


